Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-14 ICML-submission/algorithm-for-strongly-linearly-separable-data.tex
in the middle of proof of Thm 3

The idea behind the algorithm is to use $K$ copies of the \textsc{Binary
Perceptron} algorithm. Each copy corresponds to a class. In each round, each
copy makes a binary decision whether the feature vector belongs its respective
class or not. If at least one copy makes a positive prediction, the algorithm
predicts accordingly. If multiple copies make positive predictions, the
algorithm chooses one of them arbitrarily. If all copies make negative
predictions, the algorithms makes a uniformly random guess.

\begin{algorithm}[h]
\caption{\textsc{Bandit Algorithm for Strongly Linearly Separable Examples}
\label{algorithm:algorithm-for-strongly-linearly-separable-examples}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$.}
\REQUIRE{Number of rounds $T$.}
\REQUIRE{Inner product space $(V,\ip{\cdot}{\cdot})$.}
\STATE{Initialize $w_1^{(1)} = w_2^{(1)} = \dots = w_K^{(1)} = 0$}
\FOR{$t=1,2,\dots,T$}
\STATE{Observe feature vector $x_t \in V$}
\STATE{Compute $S_t = \left\{ i ~:~ 1 \le i \le K, \ \ip{w_i^{(t)}}{x_t} \ge 0 \right\}$}
\IF{$S_t = \emptyset$}
\STATE{Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$}
\STATE{Observe feedback $z_t \in \{0,1\}$ \\ \qquad where $z_t = \indicator{\widehat y_t \neq y_t}$}
\IF{$z_t = 1$}
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
\ELSE
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ \\ \qquad  for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
\STATE{Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} + x_t$}
\ENDIF
\ELSE
\STATE{Predict $\widehat y_t \in S_t$ chosen arbitrarily}
\STATE{Observe feedback $z_t \in \{0,1\}$ \\ \qquad where $z_t = \indicator{\widehat y_t \neq y_t}$}
\IF{$z_t = 1$}
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ \\ \qquad for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
\STATE{Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} - x_t$}
\ELSE
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
\ENDIF
\ENDIF
\ENDFOR
}
\end{algorithmic}
\end{algorithm}

\begin{theorem}[Mistake upper bound]
\label{theorem:strongly-separable-example-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be any inner product space, let $K$ be a positive
integer, let $\gamma$ a positive real number, let $R$ be a non-negative real number. If
$(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that the examples are strongly separable with
margin $\gamma$ and $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$ then the
expected number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes is at most $(K-1) \lfloor (R/\gamma)^2 \rfloor$.
\end{theorem}

\begin{proof}
Let $M = \sum_{t=1}^T z_t$ be the number of mistakes the algorithm makes. Let $A
= \sum_{t ~:~ S_t \neq \emptyset} z_t$ be the number of mistakes in the rounds
when $S_t \neq \emptyset$ and let $B = \sum_{t ~:~ S_t = \emptyset} z_t$ be the
number of mistakes in the rounds when $S_t = \emptyset$. Obviously, $M = A + B$.
We upper bound $A$ and $B$ separately.

Let $C$ be the number of times line 12 gets executed. Let $U$ be the number of
times line 12 and line 19 get executed. In other words, $U$ is the number of
times the $K$-tuple of vectors $(w_1^{(t)}, w_2^{(t)}, \dots, w_K^{(t)})$ gets
updated. Clearly, $U = A + C$.

@@

To finish the proof, we need to upper bound the number of updates, $U$. We claim
that $U \le \lfloor (R/\gamma)^2 \rfloor$. The proof of this upper bound is
similar to the proof of the mistake bound for \textsc{Multiclass Perceptron}
algorithm. Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors that satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2} and
\eqref{equation:strong-linear-separability-3}.
The $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots, w_K^{(t)})$
changes only if there is an update in round $t$.
We investigate how $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ and
$\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$ change. If there is an update in round $t$,
we have
\begingroup
\allowdisplaybreaks
\begin{align*}
& \sum_{i=1}^K \norm{w_i^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t}^2 \\
& = \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 + \underbrace{(-1)^{z_t} 2 \ip{w_{\widehat y_t}^{(t)}}{x_t}}_{\le 0} \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + R^2 \; .
\end{align*}
\endgroup
Therefore, after $U$ updates,
$$
\sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le R^2 U \; .
$$
Similarly, if there is an update in round $t$, we have
\begin{align*}
& \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t+1)}} \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) \\
& \qquad + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t} \\
& = \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + (-1)^{z_t} \ip{w_{\widehat y_t}^*}{x_t} \\
& \ge \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \gamma
\end{align*}
where the last inequality follows from \eqref{equation:strong-linear-separability-2}
and \eqref{equation:strong-linear-separability-3} and analysis of the cases $z_t = 0$ and $z_t = 1$.
Thus, after $U$ updates,
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}} \ge \gamma U \; .
$$
Applying Cauchy-Schwartz's inequality twice and assumption
\eqref{equation:strong-linear-separability-1}, we get that
\begin{align*}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}}
& \le \sum_{i=1}^K \norm{w_i^*} \cdot \norm{w_i^{(T+1)}} \\
& \le \sqrt{\sum_{i=1}^K \norm{w_i^*}^2} \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \\
& \le \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \; .
\end{align*}
Combining the upper and lower bounds on $\sum_{i=1}^K \norm{w_i^{(T+1)}}^2$ inequalities, we get
$$
(\gamma U)^2 \le \sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le R^2 U \; .
$$
We conclude that $U \le (R/\gamma)^2$. Since $U$ is an integer, $U \le \lfloor (R/\gamma)^2 \rfloor$.
\end{proof}

>>>>

The idea behind the algorithm is to use $K$ copies of the \textsc{Binary
Perceptron} algorithm, where each copy corresponds to a class.
In each round, copy $i$
makes a binary prediction, indicating whether the example belongs to class $i$ or not.
If at least one copy makes a positive prediction, the algorithm
chooses its prediction as one of the classes corresponding to a positive prediction.
If all copies make negative
predictions, the algorithms makes a prediction uniformly at random.

%accordingly. If multiple copies make positive predictions, the
%algorithm chooses one of them arbitrarily

\begin{algorithm}[h]
\caption{\textsc{Bandit Algorithm for Strongly Linearly Separable Examples}
\label{algorithm:algorithm-for-strongly-linearly-separable-examples}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$.}
\REQUIRE{Number of rounds $T$.}
\REQUIRE{Inner product space $(V,\ip{\cdot}{\cdot})$.}
\STATE{Initialize $w_1^{(1)} = w_2^{(1)} = \dots = w_K^{(1)} = 0$}
\FOR{$t=1,2,\dots,T$}
\STATE{Observe feature vector $x_t \in V$}
\STATE{Compute $S_t = \left\{ i ~:~ 1 \le i \le K, \ \ip{w_i^{(t)}}{x_t} \ge 0 \right\}$}
\IF{$S_t = \emptyset$}
\STATE{Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$}
\STATE{Observe feedback $z_t \in \{0,1\}$ \\ \qquad where $z_t = \indicator{\widehat y_t \neq y_t}$}
\IF{$z_t = 1$}
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
\ELSE
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ \\ \qquad  for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
\STATE{Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} + x_t$}
\label{line:pos-update}
\ENDIF
\ELSE
\STATE{Predict $\widehat y_t \in S_t$ chosen arbitrarily}
\STATE{Observe feedback $z_t \in \{0,1\}$ \\ \qquad where $z_t = \indicator{\widehat y_t \neq y_t}$}
\IF{$z_t = 1$}
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ \\ \qquad for all $i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}$}
\STATE{Update $w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} - x_t$}
\label{line:neg-update}
\ELSE
\STATE{Set $w_i^{(t+1)} = w_i^{(t)}$ for all $i \in \{1,2,\dots,K\}$}
\ENDIF
\ENDIF
\ENDFOR
}
\end{algorithmic}
\end{algorithm}

\begin{theorem}[Mistake upper bound]
\label{theorem:strongly-separable-example-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number. If
$(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that
\begin{enumerate}
  \item the examples are strongly separable with margin $\gamma$,
  \item $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$,
\end{enumerate}
then the
expected number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes is at most $(K-1) \lfloor (R/\gamma)^2 \rfloor$.
\end{theorem}

\begin{proof}
Let $M = \sum_{t=1}^T z_t$ be the number of
mistakes Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} makes.
Let $A
= \sum_{t ~:~ S_t \neq \emptyset} z_t$ be the number of mistakes in the rounds
when $S_t \neq \emptyset$ and let $B = \sum_{t ~:~ S_t = \emptyset} z_t$ be the
number of mistakes in the rounds when $S_t = \emptyset$.
It can be easily seen that $M = A + B$. %We now upper bound $A$ and $B$ separately.

Let $C$ be the number of times line~\ref{line:pos-update} gets executed. Let $U$ be the number of
times line~\ref{line:pos-update} and line~\ref{line:neg-update} get executed. In other words, $U$ is the number of
times the $K$-tuple of vectors $(w_1^{(t)}, w_2^{(t)}, \dots, w_K^{(t)})$ gets
updated. It can be easily seen that $U = A + C$.

@@

To finish the proof, we need to upper bound the number of updates $U$. We claim
that $U \le \lfloor (R/\gamma)^2 \rfloor$. The proof of this upper bound is
similar to the proof of the mistake bound for \textsc{Multiclass Perceptron}
algorithm. Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors that satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2} and
\eqref{equation:strong-linear-separability-3}.
The $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots, w_K^{(t)})$
changes only if there is an update in round $t$.
We investigate how $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ and
$\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$ change. If there is an update in round $t$,
by lines~\ref{line:pos-update} and~\ref{line:neg-update}, we always have
$ w_{\widehat y_t}^{(t+1)} = w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t $,
and for all $i \neq \widehat y_t$, $w_{i}^{(t+1)} = w_{i}^{(t)}$.
Therefore,
\begingroup
\allowdisplaybreaks
\begin{align*}
& \sum_{i=1}^K \norm{w_i^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t}^2 \\
& = \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 + \underbrace{(-1)^{z_t} 2 \ip{w_{\widehat y_t}^{(t)}}{x_t}}_{\le 0} \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + R^2 \; .
\end{align*}
\endgroup
Hence, after $U$ updates,
$$
\sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le R^2 U \; .
$$
Similarly, if there is an update in round $t$, we have
\begin{align*}
& \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t+1)}} \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) \\
& \qquad + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t} \\
& = \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + (-1)^{z_t} \ip{w_{\widehat y_t}^*}{x_t} \\
& \ge \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \gamma
\end{align*}
where the last inequality follows from \eqref{equation:strong-linear-separability-2}
and \eqref{equation:strong-linear-separability-3} and analysis of the cases $z_t = 0$ and $z_t = 1$.
Thus, after $U$ updates,
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}} \ge \gamma U \; .
$$
Applying Cauchy-Schwartz's inequality twice, and using assumption
\eqref{equation:strong-linear-separability-1}, we get that
\begin{align*}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}}
& \le \sum_{i=1}^K \norm{w_i^*} \cdot \norm{w_i^{(T+1)}} \\
& \le \sqrt{\sum_{i=1}^K \norm{w_i^*}^2} \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \\
& \le \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \; .
\end{align*}
Combining the upper and lower bounds on $\sum_{i=1}^K \norm{w_i^{(T+1)}}^2$ inequalities, we get
$$
(\gamma U)^2 \le \sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le R^2 U \; .
$$
We conclude that $U \le (R/\gamma)^2$. Since $U$ is an integer, $U \le \lfloor (R/\gamma)^2 \rfloor$.
\end{proof}

### 2019-01-14 ICML-submission/notions-of-linear-separability.tex
in the middle of proof of Thm 3

The examples are said to be \emph{strongly linearly separable with a
margin $\gamma$} if there exist vectors $w_1, w_2, \dots, w_K \in V$ such
that
\begin{align}
\label{equation:strong-linear-separability-1}
& \sum_{i=1}^K \norm{w_i}^2 \le 1 \; , \\
\label{equation:strong-linear-separability-2}
& \forall t \in \{1,2,\dots,T\} \qquad \qquad \ip{x_t}{w_{y_t}} \ge \gamma \; , \\
\label{equation:strong-linear-separability-3}
& \begin{gathered}
\forall t \in \{1,2,\dots,T\} \qquad \forall i \in \{1,2,\dots, K\} \setminus \{y_t\} \\
\ip{x_t}{w_i} \le - \gamma \; .
\end{gathered}
\end{align}
\end{definition}

@@

The notion of strong linear separability is new. Intuitively, strong linear
separability means that, for each $i$, the set of examples with class $i$ and
the set of example of other classes are separated by a linear classifier $w_i$.
Formally, $w_i$ separates the set of examples $\cbr{(x_t, y_t): t \in
\cbr{1,\ldots,T}, y_t = i}$ and $\cbr{(x_t, y_t): t \in \cbr{1,\ldots,T}, y_t
\neq i}$.

\chicheng{I think the notion has appeared in~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}, however,
they proposed an algorithm that doesn't achieve a finite mistake bound in the bandit
setting under this assumption (although they claimed that they have done so, the proof was
wrong..)}

It is easy to see that if a set of labeled examples are strongly linearly
separable with margin $\gamma$, then they are weakly linearly separable with margin
$2\gamma$. Indeed, if $w_1, w_2, \dots, w_K \in V$
satisfy \eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3} then they satisfy
\eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2} with margin $\gamma' = 2\gamma$.

Similarly, if $K=2$ and labeled examples are weakly linearly separable with a margin
$\gamma$ then they are strongly linearly separable with margin $\gamma/2$.
Indeed, if $w_1, w_2$ satisfy \eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2} then $w_1' = \frac{w_1 - w_2}{2}$, $w_2' =
\frac{w_2 - w_1}{2}$ and $\gamma'=\gamma/2$ satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3}. The last two conditions are easy
to see. The first condition follows from $\norm{w_i'}^2 \le (\frac{1}{2}
\norm{w_1} + \frac{1}{2} \norm{w_2})^2 \le \frac{1}{2}\norm{w_1}^2 +
\frac{1}{2}\norm{w_2}^2 \le \frac{1}{2}$ for $i=1,2$.

\chicheng{Shall we change the $\gamma$ in Equations
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3} to $\gamma/2$, to simplify the
presentation?}

However, for any $K \ge 3$ and any inner product space of dimension at least
$2$, there exists a set of labeled examples that is weakly linearly separable with a
positive margin $\gamma$ but it is not strongly linearly separable with any
positive margin $\gamma$. An example of such set of labeled examples is shown in
Figure~\ref{figure:weakly-linearly-separable-examples-with-margin}.

\begin{figure}
\begin{center}
\input{figures/weakly-linearly-separable-examples-with-margin}
\end{center}
\caption[]{The figure shows labeled examples in $\R^2$. The examples belong to
$K=3$ classes colored white, gray and black respectively. Each class lies in a
$120^\circ$ wedge. In other words, each class lies in an intersection of two
halfspaces.

While the examples are weakly linearly separable with a positive margin $\gamma$, they
are \emph{not} strongly linearly separable with any positive margin $\gamma$.
}
\label{figure:weakly-linearly-separable-examples-with-margin}
\end{figure}

>>>>

The examples are said to be \emph{strongly linearly separable with a
margin $\gamma$} if there exist vectors $w_1, w_2, \dots, w_K \in V$ such
that
\begin{align}
\label{equation:strong-linear-separability-1}
& \sum_{i=1}^K \norm{w_i}^2 \le 1 \; , \\
\label{equation:strong-linear-separability-2}
& \forall t \in \{1,2,\dots,T\} \qquad \qquad \ip{x_t}{w_{y_t}} \ge \frac \gamma 2 \; , \\
\label{equation:strong-linear-separability-3}
& \begin{gathered}
\forall t \in \{1,2,\dots,T\} \qquad \forall i \in \{1,2,\dots, K\} \setminus \{y_t\} \\
\ip{x_t}{w_i} \le - \frac \gamma 2 \; .
\end{gathered}
\end{align}
\end{definition}

@@

The notion of strong linear separability has appeared in the literature
(see e.g.~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}).
\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009} claimed that their
Conservative OVA algorithm with PA-I update has a finite mistake bound under the strong
linear separability assumption, their proof of Theorem 2 is problematic: their
Lemma 4 cannot be directly applied to the bandit setting,
as the algorithm only makes updates on a subset of the rows of the weight matrix.}
Intuitively, strong linear
separability means that, for each class $i$, the set of examples belonging to class $i$ and
the set of example belonging to the remaining $K-1$ classes are separated by
a linear classifier $w_i$ by a margin $\frac \gamma 2$.
%Formally, $w_i$ separates the set of examples $\cbr{(x_t, y_t): t \in
%\cbr{1,\ldots,T}, y_t = i}$ and $\cbr{(x_t, y_t): t \in \cbr{1,\ldots,T}, y_t
%\neq i}$.

%\chicheng{I think the notion has appeared in, however,
%they proposed an algorithm that doesn't achieve a finite mistake bound in the bandit
%setting under this assumption (although they claimed that they have done so, the proof was
%wrong..)}

It is easy to see that if a set of labeled examples are strongly linearly
separable with margin $\gamma$, then they are weakly linearly separable with
the same margin. Indeed, if $w_1, w_2, \dots, w_K \in V$
satisfy \eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3} then they satisfy
\eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2}.

%with margin $\gamma' = 2\gamma$.

In the special case of $K=2$, if a set of labeled examples are weakly linearly separable with a margin
$\gamma$, then they are strongly linearly separable with the same margin.
Indeed, if $w_1, w_2$ satisfy \eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2} then $w_1' = \frac{w_1 - w_2}{2}$, $w_2' =
\frac{w_2 - w_1}{2}$ satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2},
\eqref{equation:strong-linear-separability-3}.
The first condition follows from $\norm{w_i'}^2 \le (\frac{1}{2}
\norm{w_1} + \frac{1}{2} \norm{w_2})^2 \le \frac{1}{2}\norm{w_1}^2 +
\frac{1}{2}\norm{w_2}^2 \le \frac{1}{2}$ for $i=1,2$. The last two condition
follows from the fact that $w_1' - w_2' = w_1 - w_2$.
However, for any $K \ge 3$ and any inner product space of dimension at least
$2$, there exists a set of labeled examples that is weakly linearly separable with a
positive margin $\gamma$ but it is not strongly linearly separable with any
positive margin $\gamma$. An example of such set of labeled examples is shown in
Figure~\ref{figure:weakly-linearly-separable-examples-with-margin}.

%\chicheng{Shall we change the $\gamma$ in Equations
%\eqref{equation:strong-linear-separability-2},
%\eqref{equation:strong-linear-separability-3} to $\gamma/2$, to simplify the
%presentation?}

\begin{figure}
\begin{center}
\input{figures/weakly-linearly-separable-examples-with-margin}
\end{center}
\caption[]{A set of labeled examples in $\R^2$. The examples belong to
$K=3$ classes colored white, gray and black respectively. Each class lies in a
$120^\circ$ wedge. In other words, each class lies in an intersection of two
halfspaces.
\chicheng{Do we have an explicit formula for $W$ in this example? It might help
to provide it.}

While the examples are weakly linearly separable with a positive margin $\gamma$, they
are \emph{not} strongly linearly separable with any positive margin $\gamma$. For instance,
there do not exist linear separators that separates the
examples belonging to the gray class from the examples belonging to the remaining two classes.
}
\label{figure:weakly-linearly-separable-examples-with-margin}
\end{figure}
